Green Nest Material Meta-Analysis

### 2025-05-11 notebook/data_analysis.qmd
Results for meta-regressions completed

@@

# lnRR Multilevel meta-analysis

```{r}
# selecting dataset which are not flagged
dataset_lnRR<-dataset_analysis%>%
   filter(lnRR_sign != "NA" | lnRR_variance != "NA")%>%
  filter(proxy_decision == "include")

@@

@@

### Senstivity Analysis with flagged proxies

@@

```{r Types of parasite}
# This analysis only includes dataset that has a proxy relating to parasite or pathogen
dataset_lnRR_parasite <- dataset_lnRR%>%
  filter(trait_type == "Parasitic_and_pathogenic")
VCV_lnRR_parasite <- vcalc(vi = lnRR_variance,
                  cluster = paper_ID,
                  obs = Observation_ID,
                  data = dataset_lnRR_parasite,
                  rho = 0.5 # assuming that the effect sizes within the same study are correlated with rho = 0.5
                  )

 MR_lnRR_parasite <-
  metafor::rma.mv(yi = lnRR_sign, # specify lnRR as the effect size measure;
                    V = VCV_lnRR_parasite,# specify lnRR's sampling variance;
                    mods = ~ - 1 + parasite_type,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                   test = "t",
                  control = list(optimizer="optim"),
                  data = dataset_lnRR_parasite)

 # Warning: 4 rows with NAs omitted from model fitting.
 # Check what rows are these..

 # save( MR_lnRR_parasite, file = here::here("GNM/figures/ MR_lnRR_parasite.Rdata"))
```

```{r}
 MR_lnRR_parasite_plot <-
  orchard_plot(
    mod_results(
        MR_lnRR_parasite,
      mod = "parasite_type",
      group = "paper_ID",
      data = dataset_lnRR_parasite
    ),
    xlab = "")+
     scale_colour_manual(values = c("darkorange", "lightgoldenrod3", "mediumpurple4", "palegreen3")) +
  scale_fill_manual(values = c("darkorange", "lightgoldenrod3", "mediumpurple4", "palegreen3")) +
  labs(x = "",
       y = "lnRR")+
  theme(axis.text.y = element_text(angle = 90, vjust = 0.5, hjust = 1))+
  ggtitle("Parasite_type")

   MR_lnRR_parasite_plot
```

2\. Time of addition of green nest material (levels: before egg hatching, after egg hatching, continuously throughout the nesting phase).

@@

```{r}
MR_lnRR_birds <-
  metafor::rma.mv(yi = lnRR_sign, # specify lnRR as the effect size measure;
                    V = VCV_lnRR,# specify lnRR's sampling variance;
                    mods = ~ - 1 + bird_species,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                  test = "t",
                  control = list(optimizer="optim"),
                  data = dataset_lnRR)

# save(MR_lnRR_hypothesis, file = here::here("GNM/figures/MR_lnRR_hypothesis.Rdata"))
```

```{r Plot lnRR meta-regression on birds}

MR_lnRR_birds_plot <-
  orchard_plot(
    mod_results(
       MR_lnRR_birds,
      mod = "bird_species",
      group = "paper_ID",
      data = dataset_lnRR
    ), xlab ="") +
     scale_colour_manual(values = c("darkorange", "lightgoldenrod3", "mediumpurple4", "palegreen3","brown4","lightcoral","grey5")) +
  scale_fill_manual(values = c("darkorange", "lightgoldenrod3", "mediumpurple4", "palegreen3","brown4","lightcoral","grey5")) +
  labs(x = "",
       y = "lnRR")+
  theme(axis.text.y = element_text(angle = 0, vjust = 0.5, hjust = 1))+
  ggtitle("Bird Species")

 MR_lnRR_birds_plot
```

@@

```{r}
MR_lnRR_trait <-
  metafor::rma.mv(yi = lnRR_sign, # specify lnRR as the effect size measure;
                    V = lnRR_variance,# specify lnRR's sampling variance;
                    mods = ~ - 1 + trait_type,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                  test = "t",
                  control = list(optimizer="optim"),
                  data = dataset_lnRR)

# save(MR_lnRR_trait, file = here::here("GNM/figures/MR_lnRR_trait.Rdata"))
```

```{r}
MR_lnRR_trait_plot <-
  orchard_plot(
    mod_results(
        MR_lnRR_trait,
      mod = "trait_type",
      group = "paper_ID",
      data = dataset_lnRR
    ),
    xlab = "")+
     scale_colour_manual(values = c("darkorange", "lightgoldenrod3", "mediumpurple4", "palegreen3","brown4","lightcoral")) +
  scale_fill_manual(values = c("darkorange", "lightgoldenrod3", "mediumpurple4", "palegreen3","brown4","lightcoral")) +
  labs(x = "",
       y = "lnRR")+
   theme(axis.text.y = element_text(angle = 0, vjust = 0.5, hjust = 1))+
  ggtitle("Type of trait studied")

   MR_lnRR_trait_plot
```

# SMDH Multilevel meta-analysis

```{r}
# selecting dataset which are not flagged
dataset_SMDH<-dataset_analysis%>%
   filter(proxy_decision == "include"| proxy_decision == "SMDH only"|proxy_decision =="contingency table") %>% filter(SMDH_sign != "NA")

SMDH_flagged_no_missingES<- SMDH_con_flagged %>% 
   filter(!proxy_comment %in% c("use 0 as ES"))
```

### Intercept only model

We are only running these analysis for the fitness proxy labelled "included". Some proxies have been flagged and we will run them in a sensitivity analysis only.

# saving the model
# saveRDS(intercept_SMDH, file = here::here("model/intercept_SMDH.rds"))

# To calculate heterogeneity, these are the functions from orchaRd:
# # I2, CV and M
sigma2_main_model_vector <- sum(intercept_SMDH$sigma2)
I2_main_model_vector <- orchaRd::i2_ml(intercept_SMDH)
CV_main_model_vector <- orchaRd::cvh1_ml(intercept_SMDH)
M_main_model_vector <- orchaRd::m1_ml(intercept_SMDH)

### Senstivity Analysis with flagged proxies

MLMA_SMDH_VCV_flagged<- rma.mv(yi = SMDH_sign, # specify lnRR as the effect size measure;
                    V = VCV_SMDH_flagged, # specify lnRR's sampling variance-covariance matrix;
                    mods = ~ 1,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                    control = list(optimizer="optim"),
                    test = "t",
                    data = SMDH_con_flagged)

summary(MLMA_SMDH_VCV_flagged)
predict(MLMA_SMDH_VCV_flagged)

MLMA_SMDH_VCV_no_missingES<- rma.mv(yi = SMDH_sign, # specify lnRR as the effect size measure;
                    V = VCV_SMDH_no_missingES, # specify lnRR's sampling variance-covariance matrix;
                    mods = ~ 1,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                    control = list(optimizer="optim"),
                    test = "t",
                    data = SMDH_sin_missingES)

summary(MLMA_SMDH_VCV_no_missingES)
predict(MLMA_SMDH_VCV_no_missingES)

MLMA_SMDH_flagged_no_missingES<- rma.mv(yi = SMDH_sign, # specify lnRR as the effect size measure;
                    V = VCV_SMDH_flagged_no_missingES, # specify lnRR's sampling variance-covariance matrix;
                    mods = ~ 1,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                    control = list(optimizer="optim"),
                    test = "t",
                    data = SMDH_flagged_no_missingES)

summary(MLMA_SMDH_flagged_no_missingES)
predict(MLMA_SMDH_flagged_no_missingES)
```

```{r}
# MLMA_SMDH_RVE<- robust(MLMA_SMDH_VCV,
#                        cluster = paper_ID,
#                        clubSandwich = TRUE)

# summary(MLMA_SMDH_RVE)
# I2_CI <- i2_ml(MLMA_SMDH_RVE) # Relative heterogeneity
# round(I2_CI,2) # absolute heterogeneity is reported in the summary of the model
```

### Plot

```{r}
# MLMA_SMDH

MLMA_SMDH_intercept_only_plot <-
  orchard_plot(
    mod_results(
      intercept_SMDH,
      mod = "1",
      group = "paper_ID",
      data = dataset_lnRR
    ),
    xlab = "Effect Size SMDH") +
  scale_fill_manual(values="darkgreen") +
  scale_colour_manual(values="darkgreen")+
  theme(legend.direction="horizontal", legend.title = element_text(size =8),
        legend.text = element_text(size = 10), 
        axis.title = element_text(size = 15),
        axis.text.x = element_text(size = 15),
        axis.text.y = element_blank()) +
  labs(
    x = "",
    y = "SMDH"
  )+
  ggtitle("Main effect of GNM on fitness using VCV Matrix")

MLMA_SMDH_intercept_only_plot

## Unimoderator multilevel meta-regressions

#### Hypothesis

```{r}
MR_SMDH_hypothesis <-
  metafor::rma.mv(yi = SMDH_sign, # specify lnRR as the effect size measure;
                    V = VCV_SMDH,# specify lnRR's sampling variance;
                    mods = ~ - 1 + Hypothesis,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                  test = "t",
                  control = list(optimizer="optim"),
                  data = dataset_SMDH)

# save(MR_SMDH_hypothesis, file = here::here("GNM/figures/MR_lnRR_hypothesis.Rdata"))
```

```{r Plot meta-regression on hypothesis SMDH}

 MR_SMDH_hypothesis_plot <-
  orchard_plot(
    mod_results(
       MR_SMDH_hypothesis,
      mod = "Hypothesis",
      group = "paper_ID",
      data = dataset_SMDH
    ), xlab ="") +
  scale_colour_manual(values = c("darkorange", "lightgoldenrod3", "mediumpurple4", "palegreen3")) +
  scale_fill_manual(values = c("darkorange", "lightgoldenrod3", "mediumpurple4", "palegreen3")) +
  labs(x = "",
       y = "SMDH")+
  theme(axis.text.y = element_text(angle = 90, vjust = 0.5, hjust = 1))+
  ggtitle("Different Hypothesis")

  MR_SMDH_hypothesis_plot
```

#### Exploratory Analysis

```{r Types of parasite SMDH}
# This analysis only includes dataset that has a proxy relating to parasite or pathogen
dataset_SMDH_parasite <- dataset_SMDH%>%
  filter(trait_type == "Parasitic_and_pathogenic")
VCV_SMDH_parasite <- vcalc(vi = SMDH_variance,
                  cluster = paper_ID,
                  obs = Observation_ID,
                  data = dataset_SMDH_parasite,
                  rho = 0.5 # assuming that the effect sizes within the same study are correlated with rho = 0.5
                  )

 MR_SMDH_parasite <-
  metafor::rma.mv(yi = SMDH_sign, # specify lnRR as the effect size measure;
                    V = VCV_SMDH_parasite,# specify lnRR's sampling variance;
                    mods = ~ - 1 + parasite_type,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                   test = "t",
                  control = list(optimizer="optim"),
                  data = dataset_SMDH_parasite)

 # Warning: 4 rows with NAs omitted from model fitting.
 # Check what rows are these..

 # save( MR_lnRR_parasite, file = here::here("GNM/figures/ MR_lnRR_parasite.Rdata"))
```

```{r}
 MR_SMDH_parasite_plot <-
  orchard_plot(
    mod_results(
        MR_SMDH_parasite,
      mod = "parasite_type",
      group = "paper_ID",
      data = dataset_SMDH_parasite
    ),
    xlab = "")+
     scale_colour_manual(values = c("darkorange", "lightgoldenrod3", "mediumpurple4", "palegreen3")) +
  scale_fill_manual(values = c("darkorange", "lightgoldenrod3", "mediumpurple4", "palegreen3")) +
  labs(x = "",
       y = "SMDH")+
  theme(axis.text.y = element_text(angle = 90, vjust = 0.5, hjust = 1))+
  ggtitle("Parasite_type")

   MR_SMDH_parasite_plot
```

2\. Time of addition of green nest material (levels: before egg hatching, after egg hatching, continuously throughout the nesting phase).

```{r}
# This analysis only includes dataset that has a proxy relating to comaparision type
dataset_SMDH_design <- dataset_SMDH %>%
  filter(!is.na(comparision_type))%>%
  mutate(comparision_type= as.factor(comparision_type))

VCV_SMDH_design <- vcalc(vi = SMDH_variance,
                  cluster = paper_ID,
                  obs = Observation_ID,
                  data = dataset_SMDH_design,
                  rho = 0.5 # assuming that the effect sizes within the same study are correlated with rho = 0.5
                  )
MR_SMDH_design <-
  metafor::rma.mv(yi = SMDH_sign, # specify lnRR as the effect size measure;
                    V = VCV_SMDH_design ,# specify lnRR's sampling variance;
                    mods = ~ - 1 + comparision_type,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                  test = "t",
                  control = list(optimizer="optim"),
                  data = dataset_SMDH_design )

# save(MR_SMDH_design, file = here::here("GNM/figures/MR_SMDH_trait.Rdata"))
```

```{r}
 MR_SMDH_design_plot <-
  orchard_plot(
    mod_results(
        MR_SMDH_design,
      mod = "comparision_type",
      group = "paper_ID",
      data = dataset_SMDH_design
    ),
    xlab = "")+
     scale_colour_manual(values = c("darkorange", "lightgoldenrod3", "mediumpurple4", "palegreen3")) +
  scale_fill_manual(values = c("darkorange", "lightgoldenrod3", "mediumpurple4", "palegreen3")) +
  labs(x = "",
       y = "SMDH")+
  theme(axis.text.y = element_text(angle = 90, vjust = 0.5, hjust = 1))+
  ggtitle("Experimental Design")

    MR_SMDH_design_plot
```

```{r}
MR_SMDH_birds <-
  metafor::rma.mv(yi = SMDH_sign, # specify lnRR as the effect size measure;
                    V = VCV_SMDH,# specify lnRR's sampling variance;
                    mods = ~ - 1 + bird_species,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                  test = "t",
                  control = list(optimizer="optim"),
                  data = dataset_SMDH)

# save(MR_SMDH_birds, file = here::here("GNM/figures/MR_SMDH_birds.Rdata"))
```

```{r Plot SMDH meta-regression on birds}

MR_SMDH_birds_plot <-
  orchard_plot(
    mod_results(
       MR_SMDH_birds,
      mod = "bird_species",
      group = "paper_ID",
      data = dataset_SMDH
    ), xlab ="") +
     scale_colour_manual(values = c("darkorange", "lightgoldenrod3", "mediumpurple4", "palegreen3","brown4","lightcoral","grey5")) +
  scale_fill_manual(values = c("darkorange", "lightgoldenrod3", "mediumpurple4", "palegreen3","brown4","lightcoral","grey5")) +
  labs(x = "",
       y = "SMDH")+
  theme(axis.text.y = element_text(angle = 0, vjust = 0.5, hjust = 1))+
  ggtitle("Bird Species")

 MR_SMDH_birds_plot
```

```{r}
MR_SMDH_trait <-
  metafor::rma.mv(yi = SMDH_sign, # specify lnRR as the effect size measure;
                    V = SMDH_variance,# specify lnRR's sampling variance;
                    mods = ~ - 1 + trait_type,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                  test = "t",
                  control = list(optimizer="optim"),
                  data = dataset_SMDH)

# save(MR_SMDH_trait, file = here::here("GNM/figures/MR_SMDH_trait.Rdata"))
```

```{r}
MR_SMDH_trait_plot <-
  orchard_plot(
    mod_results(
        MR_SMDH_trait,
      mod = "trait_type",
      group = "paper_ID",
      data = dataset_SMDH
    ),
    xlab = "")+
     scale_colour_manual(values = c("darkorange", "lightgoldenrod3", "mediumpurple4", "palegreen3","brown4","lightcoral")) +
  scale_fill_manual(values = c("darkorange", "lightgoldenrod3", "mediumpurple4", "palegreen3","brown4","lightcoral")) +
  labs(x = "",
       y = "SMDH")+
   theme(axis.text.y = element_text(angle = 0, vjust = 0.5, hjust = 1))+
  ggtitle("Type of trait studied")

   MR_SMDH_trait_plot
```

@@

1.  We will re-run all our meta-analyses and meta-regressions only using effect sizes calculated from means, SDs, and ns (i.e., excluding any effect sizes calculated from inferential statistics)

2.  For the overall effect size (i.e., the meta-analytical mean), we will run the bivariate multilevel meta-analytic model suggested by Yang et al. (2024), which models both lnRR and SMD(H) simultaneously (i.e., including them both as response variables and estimating their correlation; random).

```{r}
dataset_bivariate <- data.frame(paper_ID = c(dataset_lnRR$paper_ID, dataset_SMDH$paper_ID),
                      Observation_ID = c(dataset_lnRR$Observation_ID, dataset_SMDH$Observation_ID),
                      ES = c(dataset_lnRR$lnRR_sign, dataset_SMDH$SMDH_sign),
                      V_ES = c(dataset_lnRR$lnRR_variance, dataset_SMDH$SMDH_variance),
                      ES_measure = c(rep("lnRR", nrow(dataset_lnRR)), rep("SMD", nrow(dataset_SMDH)))
                      )

VCV <- vcalc(vi = V_ES, # sampling variances of lnRR and SMD that are correlated within the same study;
             cluster = paper_ID, # study identity;
             type = ES_measure, # different types of effect size measures underlying the observed effect sizes;
             data = dataset_bivariate, # the long format data frame;
             rho = 0.5, # assuming that the effect sizes within the same study are correlated with rho = 0.5.
             nearpd = TRUE # specify whether impose the non positive definite VCV matrix (not invertible) to the nearest positive semi-definite matrix 
             )

>>>>

## Loading Data for analysis

@@

# Preparing datasets

```{r seperating datasets for analysis}

## For lnRR
# selecting dataset which are not flagged
dataset_lnRR<-dataset_analysis%>%
   filter(lnRR_sign != "NA" | lnRR_variance != "NA")%>%
  filter(proxy_decision == "include")

# selecting only published dataset
dataset_lnRR_published<-dataset_lnRR%>%
  filter(!is.na(year_publication))

@@

# selecting dataset without those using 0 as ES because authors reported no effect
lnRR_sin_missingES <- dataset_lnRR %>% 
   filter(!proxy_comment %in% c("use 0 as ES"))

## For SMDH
# selecting dataset which are not flagged
dataset_SMDH<-dataset_analysis%>%
   filter(proxy_decision == "include"| proxy_decision == "SMDH only"|proxy_decision =="contingency table") %>% filter(SMDH_sign != "NA")

# selecting only published dataset
dataset_SMDH_published<-dataset_SMDH%>%
  filter(!is.na(year_publication))

@@

## using SMDH

We are only running these analysis for the fitness proxy labelled "included". Some proxies have been flagged and we will run them in a sensitivity analysis only

```{r intercept-only-model-SMDH}

# saving the model
# saveRDS(intercept_SMDH, file = here::here("model/intercept_SMDH.rds"))
```

## SMDH-heterogeneity

# To calculate heterogeneity, these are the functions from orchaRd:
# # I2, CV and M
sigma2_main_model_vector <- sum(intercept_SMDH$sigma2)
# 0.54, 1.25, and 3.03 for SMD;
I2_main_model_vector <- orchaRd::i2_ml(intercept_SMDH)
# 0.78, 0.89, and 0.96 for SMD
# 0.88, 0.95, and 0.99 for lnRR
CV_main_model_vector <- orchaRd::cvh1_ml(intercept_SMDH)
# CVH2 (CVH1) - 1.36 (1.16), 3.76 (1.94), and 12.1 (3.48) for lnRR
M_main_model_vector <- orchaRd::m1_ml(intercept_SMDH)
# M2 (M1) - 0.58 (0.54), 0.79 (0.66), and 0.78 for lnRR

## Bivariate Meta-analytical model

For the overall effect size (i.e., the meta-analytical mean), we will run the bivariate multilevel meta-analytic model suggested by Yang et al. (2024), which models both lnRR and SMD(H) simultaneously (i.e., including them both as response variables and estimating their correlation; random).

```{r}
dataset_bivariate <- data.frame(paper_ID = c(dataset_lnRR$paper_ID, dataset_SMDH$paper_ID),
                      Observation_ID = c(dataset_lnRR$Observation_ID, dataset_SMDH$Observation_ID),
                      group_ID_coded = c(dataset_lnRR$group_ID_coded, dataset_SMDH$group_ID_coded),
                      population_ID = c(dataset_lnRR$population_ID, dataset_SMDH$population_ID),
                      bird_species = c(dataset_lnRR$bird_species, dataset_SMDH$bird_species),
                      ES = c(dataset_lnRR$lnRR_sign, dataset_SMDH$SMDH_sign),
                      V_ES = c(dataset_lnRR$lnRR_variance, dataset_SMDH$SMDH_variance),
                      ES_measure = c(rep("lnRR", nrow(dataset_lnRR)), rep("SMDH", nrow(dataset_SMDH)))
                      )

VCV_bivariate <- vcalc(vi = V_ES, # sampling variances of lnRR and SMD that are correlated within the same study;
             cluster = paper_ID, # study identity;
             type = ES_measure, # different types of effect size measures underlying the observed effect sizes;
             data = dataset_bivariate, # the long format data frame;
             rho = 0.2, # assuming that the effect sizes within the same study are correlated with rho = 0.5.
             nearpd = TRUE # specify whether impose the non positive definite VCV matrix (not invertible) to the nearest positive semi-definite matrix 
             )

Bivariate_model <- rma.mv(yi = ES, # specify the effect size estimate (the variable ES in our case);
                V = VCV_bivariate, # specify the imputed variance-covariance matrix;
                mods = ~ ES_measure - 1, # specify the variable "ES_measure" indicating the types of effect size measures;
                random = list(~ ES_measure | paper_ID, 
                              ~ ES_measure | Observation_ID,
                              ~ 1 | group_ID_coded,
                              ~ 1 | population_ID,
                              ~ 1 | bird_species
                              ), # add correlated random effects corresponding to the lnRR and SMD parameters in the same study;
                struct = "UN", # impose an unstructured variance-covariance of the study-specific random effects;
                test = "t", # t distribution is specified to test the overall effect against the null hypothesis and construct confidence intervals;
                method = "REML", # restricted likelihood maximum is assigned as the estimator for variance components as suggested;
                data = dataset_bivariate, # the long format dataset
                sparse = T
               )

saveRDS(Bivariate_model, file = here::here("sensitivity_analysis/bivariate_model.rds"))

summary(Bivariate_model)
```

```{r WARNING = F}
pacman::p_load(ellipse)
ab.l <- matreg(y=2, x=1, R=Bivariate_model$G, cov=TRUE, means=coef(Bivariate_model), n=Bivariate_model$g.levels.comb.k) # fit regression model

xy <- ellipse(Bivariate_model$G, centre=coef(Bivariate_model), level=0.95) # get 95% coverage region
ellipse_df <- data.frame(x = xy[, 1], y = xy[, 2]) # convert it int dataframe so that we can use ggplot to make a figure

pivot_wider(dataset_bivariate, names_from = ES_measure, values_from = c(ES, V_ES)) %>%
  ggplot() + 
  geom_point(aes(x = ES_lnRR, y = ES_SMDH, size = 1/sqrt(V_ES_lnRR)), color = "#1B9E77", alpha = 0.5) + 
  geom_point(aes(x = coef(Bivariate_model)[1], y = coef(Bivariate_model)[2]), color = "red", size = 2) + 
  geom_abline(intercept = ab.l$tab$beta[1], slope = ab.l$tab$beta[2]) +
  geom_path(data = ellipse_df, aes(x = x, y = y), color = "gray50") +
  #scale_x_continuous(limits = c(-2.5, 2.5)) + 
  #scale_y_continuous(limits = c(-2.5, 2.5)) + 
  guides(size = "none") + 
  labs(x = "lnRR", y = "SMD") + 
  theme_bw()
```

There is a very high negative correlation between the two effect sizes. This probably is an artifact of the data and not a true correlation. If you look at the model summary, the $tau^2$ for lnRR, it is extremely small 0.0008, i.e., there is very little variance. This could also be stemming from the fact that the log values in the dataset could be causing this. For SMDH on the other hand, the $tau^2$ is 0.1681 that is much larger. This difference in the variance estimate itself could be causing this artifact of the model. I will only go ahead to present this model as a senstivity analysis because I am not completely convinced by it's robustness.

# Sensitivity Analysis

### Sensitivity Analysis with published data only

VCV_lnRR_published <- vcalc(vi = lnRR_variance,
                  cluster = paper_ID,
                  obs = Observation_ID,
                  data = dataset_lnRR_published,
                  rho = 0.5 # assuming that the effect sizes within the same study are correlated with rho = 0.5
                  )

SA_lnRR_published<- rma.mv(yi = lnRR_sign, # specify lnRR as the effect size measure;
                    V = VCV_lnRR_published, # specify lnRR's sampling variance-covariance matrix;
                    mods = ~ 1,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                    control = list(optimizer="optim"),
                    test = "t",
                    data = dataset_lnRR_published)

# saving the model
saveRDS(SA_lnRR_published, file = here::here("sensitivity_analysis/SA_lnRR_published.rds"))

summary(SA_lnRR_published)
predict(SA_lnRR_published)

VCV_SMDH_published <- vcalc(vi = SMDH_variance,
                  cluster = paper_ID,
                  obs = Observation_ID,
                  data = dataset_SMDH_published,
                  rho = 0.5 # assuming that the effect sizes within the same study are correlated with rho = 0.5
                  )

SA_SMDH_published<- rma.mv(yi = SMDH_sign, # specify lnRR as the effect size measure;
                    V = VCV_SMDH_published, # specify lnRR's sampling variance-covariance matrix;
                    mods = ~ 1,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                    control = list(optimizer="optim"),
                    test = "t",
                    data = dataset_SMDH_published)

# saving the model
saveRDS(SA_SMDH_published, file = here::here("sensitivity_analysis/SA_SMDH_published.rds"))

summary(SA_SMDH_published)
predict(SA_SMDH_published)

### Sensitivity Analysis with flagged proxies

@@

```{r Types of parasite}
# This analysis only includes dataset that has a proxy relating to parasite or pathogen
dataset_lnRR_parasite <- dataset_lnRR%>%
  filter(!is.na(parasite_type))%>%
  mutate(parasite_type=as.factor(parasite_type),
         paper_ID=as.factor(paper_ID),
         Observation_ID=as.factor(Observation_ID),
         group_ID_coded=as.factor(group_ID_coded),
         bird_species=as.factor(bird_species))

VCV_lnRR_parasite <- vcalc(vi = lnRR_variance,
                  cluster = paper_ID,
                  obs = Observation_ID,
                  data = dataset_lnRR_parasite,
                  rho = 0.5 # assuming that the effect sizes within the same study are correlated with rho = 0.5
                  )

lnRR_parasite <-
  metafor::rma.mv(yi = lnRR_sign, # specify lnRR as the effect size measure;
                    V = VCV_lnRR_parasite,# specify lnRR's sampling variance;
                    mods = ~ - 1 + parasite_type,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                   test = "t",
                  control = list(optimizer="optim"),
                  data = dataset_lnRR_parasite)

saveRDS( lnRR_parasite, file = here::here("model/lnRR_parasite.rds"))

### Similiarly for SMDH as well selecting the dataset with parasite type
# This analysis only includes dataset that has a proxy relating to parasite or pathogen
dataset_SMDH_parasite <- dataset_SMDH%>%
  filter(!is.na(parasite_type))
VCV_SMDH_parasite <- vcalc(vi = SMDH_variance,
                  cluster = paper_ID,
                  obs = Observation_ID,
                  data = dataset_SMDH_parasite,
                  rho = 0.5 # assuming that the effect sizes within the same study are correlated with rho = 0.5
                  )

SMDH_parasite <-
  metafor::rma.mv(yi = SMDH_sign, # specify lnRR as the effect size measure;
                    V = VCV_SMDH_parasite,# specify lnRR's sampling variance;
                    mods = ~ - 1 + parasite_type,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                   test = "t",
                  control = list(optimizer="optim"),
                  data = dataset_SMDH_parasite)

saveRDS(SMDH_parasite, file = here::here("model/SMDH_parasite.rds"))
```

#### Time of addition of green nest material

(levels: before egg hatching, after egg hatching, continuously throughout the nesting phase).

```{r}
## We use the complete dataset in this case
lnRR_time_gnm_addition <-
  metafor::rma.mv(yi = lnRR_sign, # specify lnRR as the effect size measure;
                    V = VCV_lnRR,# specify lnRR's sampling variance using the variance-covariance matrix we created for the intercept-only model since the whole dataset is being used in this case as well
                    mods = ~ - 1 + time_of_gnm_addition,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                  test = "t",
                  control = list(optimizer="optim"),
                  data = dataset_lnRR)

saveRDS(lnRR_time_gnm_addition, file = here::here("model/lnRR_time_gnm_addition.rds"))

## Now similarly for the SMDH effect size using the complete dataset
SMDH_time_gnm_addition <-
  metafor::rma.mv(yi = SMDH_sign, # specify lnRR as the effect size measure;
                    V = VCV_SMDH,# specify lnRR's sampling variance using the variance-covariance matrix we created for the intercept-only model since the whole dataset is being used in this case as well
                    mods = ~ - 1 + time_of_gnm_addition,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                  test = "t",
                  control = list(optimizer="optim"),
                  data = dataset_SMDH)

saveRDS(SMDH_time_gnm_addition, file = here::here("model/SMDH_time_gnm_addition.rds"))
```

@@

```{r}
lnRR_birds <-
  metafor::rma.mv(yi = lnRR_sign, # specify lnRR as the effect size measure;
                    V = VCV_lnRR,# specify lnRR's sampling variance;
                    mods = ~ - 1 + bird_species,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                  test = "t",
                  control = list(optimizer="optim"),
                  data = dataset_lnRR)

saveRDS(lnRR_birds, file = here::here("model/lnRR_birds.rds"))

## Now doing this for SMDH

SMDH_birds <-
  metafor::rma.mv(yi = SMDH_sign, # specify lnRR as the effect size measure;
                    V = VCV_SMDH,# specify lnRR's sampling variance;
                    mods = ~ - 1 + bird_species,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                  test = "t",
                  control = list(optimizer="optim"),
                  data = dataset_SMDH)

saveRDS(SMDH_birds, file = here::here("model/SMDH_birds.rds"))
```

@@

```{r}
lnRR_trait <-
  metafor::rma.mv(yi = lnRR_sign, # specify lnRR as the effect size measure;
                    V = VCV_lnRR,# specify lnRR's sampling variance;
                    mods = ~ - 1 + trait_type,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                  test = "t",
                  control = list(optimizer="optim"),
                  data = dataset_lnRR)

saveRDS(lnRR_trait, file = here::here("model/lnRR_trait.rds"))

## Now similarly for SMDH

SMDH_trait <-
  metafor::rma.mv(yi = SMDH_sign, # specify lnRR as the effect size measure;
                    V = VCV_SMDH,# specify lnRR's sampling variance;
                    mods = ~ - 1 + trait_type,
                    random = list(~ 1 | paper_ID,
                                  ~ 1 | Observation_ID,
                                  ~ 1 | group_ID_coded,
                                  ~ 1 | population_ID,
                                  ~ 1 | bird_species),
                    method = "REML",
                  test = "t",
                  control = list(optimizer="optim"),
                  data = dataset_SMDH)

saveRDS(SMDH_trait, file = here::here("model/SMDH_trait.rds"))
```

@@
